---
title: "Multiple Factor Analysis (MFA) with FactoMineR on the sensory description of the wines"
author: "Francois Husson"
output:
  pdf_document:
    fig_height: 6
    keep_tex: yes
  html_document: default
  word_document: default
---


# Loading FactoMineR and data
```{r}
library(FactoMineR)
data(wine)
```

# MFA
```{r}
res <- MFA(wine, group=c(2,5,3,10,9,2), type=c("n",rep("s",5)),
    ncp=5, name.group=c("orig","smell","vis","smellAf","tasting","pref"),
    num.group.sup=c(1,6))
```

# A summary of the main results with the summary.MFA function
The ```summary``` function gives the main results of the MFA. 

```{r eval=FALSE}
summary(res)
```

We just want the results on the first 2 dimensions (by default, the function returns the first 3 dimensions).

```{r}
summary(res, ncp=2)
```

# Graph of the individuals and the categories
```{r}
plot(res)
plot(res, invisible="quali")
plot(res, invisible="quali", cex=0.8)
plot(res, invisible="quali", cex=0.8, partial=c("1VAU","PER1"))
```

# Selection in the graph of the individuals and categories
```{r}
plot(res, invisible="quali", cex=0.8, partial=c("1VAU","PER1"), select="cos2 0.4")
plot(res, ,invisible="quali", partial="all", 
     palette=palette(c("black","transparent","transparent","blue","transparent")))
plot(res, invisible="quali", habillage=1, cex=0.8, select="cos2 0.4")
plot(res, invisible="quali", habillage="Soil", cex=0.8, select="cos2 0.4")
plot(res, invisible="quali", habillage="Soil", cex=0.8, select="contrib 5")
plot(res, invisible="quali", habillage="Soil", cex=0.8, select="contrib 5", unselect=0)
plot(res, invisible="quali", habillage="Soil", cex=0.8, select="contrib 5", unselect=1)
plot(res, invisible="quali", habillage="Soil", cex=0.8, select="contrib 5", unselect="grey70")
plot(res, invisible="quali", habillage="Soil", cex=0.8, select="contrib 5", 
     unselect="grey70", axes=3:4)
plot(res, invisible="quali", habillage="Soil", cex=0.8, select=c("1VAU","PER1"), 
     unselect="grey70", axes=3:4)
```

# Graph of the variables
```{r}
plot(res, choix="var")
plot(res, choix="var", shadow=TRUE)
plot(res, choix="var", shadow=TRUE, select="contrib 5")
plot(res, choix="var", shadow=TRUE, select="contrib 5", axes=3:4)
```

# Graph of the partial axes
```{r}
res <- MFA(wine, group=c(2,5,3,10,9,2), type=c("n",rep("s",5)),
    ncp=3, name.group=c("orig","smell","vis","smellAf","tasting","pref"),
    num.group.sup=c(1,6))
plot(res, choix="axes")
```

# Description of dimensions
```{r}
dimdesc(res)
```

