---
title: "Hierarchical ascendant clustering with FactoMineR (Temperature example)"
author: "Magalie Houée-Bigot & François Husson"
output:
  pdf_document:
    fig_height: 6
    keep_tex: yes
  word_document: default
---



### Import data 
```{r}
setwd("C:/Users/houee/Downloads") # select the working directory 
temperature <- read.table("data_clustering_Temperature.csv",
        header=TRUE, sep=";", dec=".", row.names=1)
```

```header=TRUE``` : indicates that the file contains the names of the variables

```sep=";"``` : indicates the fields separator (usually ";" or "," for csv files)

```row.names=1``` : indicates the column of the table which contains the row names

It is important to check that the import is well done


```{r}
summary(temperature)
```

# Loading FactoMineR
```{r}
library(FactoMineR)
```

# PCA
```{r eval=FALSE}
res <- PCA(temperature, ind.sup=24:35, quanti.sup=13:16,quali.sup=17, ncp=8)
```
```{r echo=FALSE}
res <- PCA(temperature, ind.sup=24:35, quanti.sup=13:16,quali.sup=17, ncp=8,graph=FALSE)
plot(res,cex=.7)
plot(res,choix="var",cex=.7)
```

# Hierarchical ascendant clustering
```{r echo=FALSE}
res.hcpc <- HCPC(res,  graph=FALSE, nb.clust=-1)
```

```{r eval=FALSE}
res.hcpc <- HCPC(res)
```

```{r echo=FALSE}
plot(res.hcpc,choice="tree")
```
```{r echo=FALSE}
plot(res.hcpc,choice="3D.map")
```
```{r echo=FALSE}
plot(res.hcpc,choice="map",draw.tree=FALSE)
```

# Outputs
```{r}
names(res.hcpc)
res.hcpc$data.clust
res.hcpc$desc.var
res.hcpc$desc.axes
res.hcpc$desc.ind
```

# Hierarchical ascendant clustering without consolidation
```{r echo=FALSE}
res.hcpc <- HCPC(res, consol=FALSE, graph=FALSE)
```

```{r eval=FALSE}
res.hcpc <- HCPC(res, consol=FALSE)
```

```{r echo=FALSE}
plot(res.hcpc,choice="tree")
```
```{r echo=FALSE}
plot(res.hcpc,choice="3D.map")
```
```{r echo=FALSE}
plot(res.hcpc,choice="map",draw.tree=FALSE)
```
