---
title: "Multiple Correspondence Analysis (MCA) with FactoMineR (tea dataset)"
author: "Magalie Houée-Bigot & François Husson"
output:
  pdf_document:
    fig_height: 6
    keep_tex: yes
  html_document: default
  word_document: default
---


# Import data 

```{r}
setwd("C:/Users/houee/Downloads") # select the working directory
tea <- read.table("data_MCA_Tea.csv", header=TRUE, sep=";")
```


```header=TRUE``` : indicates that the file contains the names of the variables

```sep=";"``` : indicates the fields separator (usually ";" or "," for csv files)

It is important to check that the import is well done

```{r}
summary(tea)
```

# Loading FactoMineR
```{r}
library(FactoMineR)
```

# MCA
```{r}
res <- MCA(tea, quanti.sup=19, quali.sup=c(20:36))
```

Outputs can be summarized with the function ```summary```. 

```{r eval=FALSE}
summary(res, ncp=3, nbelements=Inf)    ## fonction summary.MCA
```

Outputs are given for the first 2 dimensions (by default 3 dimensions are given) and for the 5 first elements.

```{r}
summary(res, ncp=2, nbelements=5)
```

# Graph of the eigenvalues
```{r}
barplot(res$eig[,2],main="Eigenvalues", names.arg=1:nrow(res$eig))
```

# Graph with some labels
```{r}
plot(res, label=c("var","quali.sup"), cex=0.7)
```

# Graphs with selection of elements
```{r}
plot(res, invisible=c("var","quali.sup"), cex=0.7)
plot(res, invisible=c("ind","quali.sup"),autoLab="y",cex=0.7,
     title="Active categories")
plot(res, invisible=c("ind","var"),autoLab="y",cex=0.7,
     title="Supplementary categories")
```

# Selection of some categories
```{r}
plot(res, invisible="ind",autoLab="y",cex=0.7,selectMod="cos2 10")
plot(res, invisible="ind",autoLab="y",cex=0.7,selectMod="contrib 20")
```

# Selection of some individuals
```{r}
plot(res, invisible=c("var","quali.sup"),autoLab="y",cex=0.7,select="cos2 20")
```

# Selection of some categories and some individuals
```{r}
plot(res, autoLab="y",cex=0.7, select="cos2 20", selectMod="cos2 10")
```

# Graphs of the variables
```{r}
plot(res, choix="var",xlim=c(0,0.6),ylim=c(0,0.6))
plot(res, choix="var",xlim=c(0,0.6),ylim=c(0,0.6),invisible=c("quali.sup","quanti.sup"))
```

# Graphs on dimensions 3 and 4
```{r}
plot(res,invisible=c("var","quali.sup"),cex=0.7,select="contrib 20",axes=3:4)
plot(res, invisible="ind",autoLab="y",cex=0.7,selectMod="cos2 20",axes=3:4)
```

# Confidence ellipses around the categories for the variables 14 to 17
```{r}
plotellipses(res,keepvar=c(14:17))
```

# Description of the dimensions
```{r}
dimdesc(res)
```
