---
title: "Multiple Correspondence Analysis (MCA) with FactoMineR (hobbies dataset)"
author: "Magalie Houée-Bigot & François Husson"
output:
  pdf_document:
    fig_height: 6
    keep_tex: yes
  html_document: default
  word_document: default
---


# Import data 

```{r}
setwd("C:/users/houee/Downloads") # select the working directory
hobbies = read.table("data_MCA_Hobbies.csv", header=TRUE, sep=";")
```

```header=TRUE``` : indicates that the file contains the names of the variables

```sep=";"``` : indicates the fields separator (usually ";" or "," for csv files)

It is important to check that the import is well done

```{r}
summary(hobbies)
```

# Transform the TV variable as factor
```{r}
hobbies[,"TV"] = as.factor(hobbies[,"TV"])
summary(hobbies[,"TV"])
```

# Loading FactoMineR
```{r}
library(FactoMineR)
```

# MCA
```{r}
res.mca <- MCA(hobbies,quali.sup=19:22,quanti.sup=23)
```

Outputs can be summarized with the function ```summary```. 

```{r eval=FALSE}
summary(res.mca)
```

Outputs are given for the first 2 dimensions (by default 3 dimensions are given).

```{r}
summary(res.mca, ncp=2)
```

# Graph of the eigenvalues
```{r}
barplot(res.mca$eig[,2],main="Eigenvalues", names.arg=1:nrow(res.mca$eig))
```

# Graphs of the individuals
```{r}
plot(res.mca,invisible=c("var","quali.sup"),cex=.5,label="none",
     title="Graph of the individuals") 
plot(res.mca,invisible=c("var","quali.sup"),cex=.5,label="none",
     title="Graph of the individuals", habillage="Gardening") 
```

# Graphs of the categories
```{r}
plot(res.mca,invis=c("ind","quali.sup"),col.var=c(rep(c("black","red"),17),"black",rep("red",4)),
     title="Graph of the active categories")
plot(res.mca,invisible=c("ind","var"),hab="quali", 
     palette=palette(c("blue","maroon","darkgreen","black","red")), 
     title="Graph of the supplementary categories")
```

# Graphs of the variables
```{r}
plot(res.mca,choix="var",title="Graph of the variables")
plot(res.mca,choix="quanti.sup",title="Graph of the continuous variables")
```

# Description of the dimensions
```{r}
dimdesc(res.mca)
```

# Confidence ellipses around the categories for the first 4 variables
```{r}
plotellipses(res.mca, cex=0.2, magnify=12, keepvar=1:4)
```
