---
title: "Principal Component Analysis (PCA) with FactoMineR (Wine dataset)"
author: "Magalie Houée-Bigot & François Husson"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Import data 

Upload the Expert Wine dataset on your computer. 
```{r eval=TRUE, echo=FALSE}
 setwd("C:/users/houee/AOBOX/MOOC_anado_GB/Données")
 wine <- read.table("data_PCA_ExpertWine.csv",header=TRUE, sep=";", dec=".", row.names=1)
```
```{r eval=FALSE, echo=TRUE}
 setwd("C:/users/houee/") # select the current directory 
 wine <- read.table("data_PCA_ExpertWine.csv",header=TRUE, sep=";", dec=".", row.names=1)
```

```header=TRUE``` : indicates that the file contains the names of the variables

```sep=";"``` : indicates the fields separator (usually ";" or "," for csv files)

```row.names=1``` : indicates the column of the table which contains the row names

It is important to check that the import is well done

```{r}
 summary(wine)
```

# Loading FactoMineR
```{r}
library(FactoMineR)
```

# PCA with supplementary variables
```{r}
res <- PCA(wine,quali.sup=1,quanti.sup=29:30,graph=FALSE)
par(cex=0.8,cex.main=1,cex.lab=0.8,cex.axis=0.8) # change graphical parameters (font size)
plot(res,choix="ind") # individuals factor map
plot(res,choix="var") # variables factor map
```

# Summary of outputs 
Outputs can be summarized with the function ```summary```.
```{r}
 summary(res, nbelements=Inf)  ## to print all the elements
```

# Bar chart of eigenvalues
```{r}
 barplot(res$eig[,1],main="Eigenvalues",names.arg=1:nrow(res$eig))
```

# Drawing wines according to the label 
```{r}
 plot(res,habillage="Label")
```