---
title: "Principal Component Analysis (PCA) with FactoMineR (decathlon dataset)"
author: "François Husson & Magalie Houée-Bigot"
output:
  pdf_document:
    fig_height: 6
    keep_tex: yes
  html_document: default
  word_document: default
---
```{r}

```


# Import data (data are imported from internet)

```{r eval=FALSE}
decathlon <- read.table("http://www.agrocampus-ouest.fr/math/RforStat/decathlon.csv", 
      header=TRUE, sep=";", row.names=1, check.names=FALSE)
```
```{r echo=FALSE}
decathlon <- read.table("http://www.agrocampus-ouest.fr/math/RforStat/decathlon.csv", 
                        header=TRUE, sep=";", row.names=1, check.names=FALSE)
```
```header=TRUE``` : indicates that the file contains the names of the variables

```sep=";"``` : indicates the fields separator (usually ";" or "," for csv files)

```row.names=1``` : indicates the column of the table which contains the row names

```check.names=FALSE``` : indicated that the names of the variables in the data frame are unchecked

It is important to check that the import is well done

```{r}
summary(decathlon)
```

# Loading FactoMineR
```{r}
library(FactoMineR)
```

# PCA with only active elements as active
```{r}
res <- PCA(decathlon[,1:10])
```

Outputs can be summarized with the function ```summary```. 

```{r eval=FALSE}
summary(res)
```

Outputs are given for the first 2 dimensions (by default 3 dimensions are given).

```{r}
summary(res, ncp=2)
```

# PCA with supplementary variables
```{r}
res <- PCA(decathlon, quanti.sup=11:12, quali.sup=13)
summary(res, ncp=2, nbelements=Inf)
```
In order to print the results in a file:
```{r eval=FALSE}
summary(res, nbelements=Inf, file="summaryResult.txt")
```

# Description of the dimensions
```{r}
dimdesc(res)
dimdesc(res, proba=0.2) # change the significance threshold to characterize the dimension
```

# Graph of the individuals with a title and a smaller size for the labels
```{r}
plot(res, cex=0.8, invisible="quali", title="Graph of the individuals")
```

With many individuals and if the labels are not useful, one can suppress the labels with the argument label="none".

```{r eval=FALSE}
plot(res, cex=0.8, invisible="quali", label="none", title="Graph of the individuals")
```

# Drawing individuals according to the competition
```{r}
plot(res, cex=0.8, habillage="Competition")
```

We could have written:
```{r eval=FALSE}
plot(res, cex=0.8, habillage=13)
```
\newpage

# Confidence ellipses around the categories
```{r}
plotellipses(res)
```

If several qualitative variables are available, there will be as many graphs as qualitative variables. And on each graph the confidence ellipses around the categories of a categorical variable.

\newpage

# Graph for dimensions 3 and 4
```{r}
plot(res, choix="ind", cex=0.8, habillage=13, title="Graph of the individuals", axes=3:4)
plot(res, choix="var", title="Graph of the variables", axes=3:4)
```

\newpage

# Selecting individuals

```select="cos2 0.7"``` : select the individuals that have a quality of representation on the map greater than 0.7

```select="cos2 5"``` : select the 5 individuals that have the best quality of representation on the map

```select="contrib 5"``` : select the 5 individuals that contribute the most to the construction of the map

```select=c("nom1","nom2")``` : select the individuals by their name

```{r}
plot(res, cex=0.8, habillage=13, select="cos2 0.7")
plot(res, cex=0.8, habillage=13, select="contrib 5")
plot(res, cex=0.8, habillage=13, select=c("Clay","Karpov"))
```

# Selecting variables
```{r}
plot(res, choix="var", select="contrib 5")
```

\newpage

# Graph with different options
```{r}
plot(res, cex=0.8, habillage=13, select="cos2 0.7", title="Decathlon", 
     cex.main=1.1, cex.axis=0.9, shadow=TRUE, auto="y")
```